---
title: "Lesson 1b - Factors"
author: "ZJ"
date: "2 October, 2014"
output: ioslides_presentation
---

## The factor data-type

We have studied

- numerics (integers and real numbers)
- boolean (true/false)
- strings (letters, and other characters and symbols)

Factors is a data type that is unique to R

- A factor is a finite set of values
- E.g. Male, Female can be the distinct values for a factor
```{r}
mf <- factor(c("Male","Female"))
mf
# levels of a factor refers to the possible distinct values
levels(mf)
```

## Factors - continued

- The *levels* are the external presentation of the factors
- Internally it's stored as integers (this fact is significant and can lead to some elegant code)
```{r}
mf
as.numeric(mf)
```

## Factors - continued
Factors can be ordinal or non-ordinal
- **Ordinal** there is some natural order structure to it e.g. Income Brackets
- **non-ordinal** - no necessary order to it e.g. apple, oranges, banana
  
## Factors - Basic summary

```{r}
summary(iris$Species)
```

## Factors - Basic summary

```{r}
plot(iris$Species)
```

## Factors - Basic summary

```{r}
table(iris$Species)
```