---
title: "Lists"
author: "ZJ"
date: "3 October, 2014"
output: ioslides_presentation
---

## List is THE R data structure

```{r}
a.list <- NULL
a.list$element1 <- 1
a.list$element2 <- c(1,2,3)
a.list
# equivalently
a.list <- list(element1 = 1, element2 = c(1,2,3))
```

## List

- Each list can have many child elements
- Child Elements are access by the \$ E.g. list.name\$child.element
- The child element can be of any type. E.g. you can have a list of lists
- The child elements don't have to be the same type of things
- Useful for grouping related information together

## Funtions you can apply to lists

```{r}
# the structure of the list
str(a.list)
# the child element's names
names(a.list)
```

## Accessing Elements

```{r}
# access child elements by string name 
# [ ] gives a list
a.list["element1"] 
a.list[c("element1","element2")] # gives a list
# [[ ]] gives a vector
a.list[["element1"]] 
```

## Exercise

1. Create a list named dog, with 3 child elements 
- weight.kg = 3, 
- age.yrs = 2, 
- name = "Lassie"

2. Inspecting the result of a linear model fit using str
```{r}
cars.linear.model <- lm(dist ~ speed, data = cars) # this builds a linear model
str(cars.linear.model)
```

## Exercise (Solution)

```{r}
# Create a list named dog, 
# with 3 child elements 
# weight.kg = 3, 
# age.yrs = 2, 
# name = "Lassie"
dog <- list(weight.kg = 3, age.yrs = 2, name = "Lassie")
```