---
title: "Lesson 3 - The ply family of functions"
author: "ZJ"
date: "3 October, 2014"
output: ioslides_presentation
---

## apply & and ply-family of functions

- The apply family of functions is important in the R world
- Most of what they can do can also be done with for/while loops 
- But apply is much faster
- Using the ply-family of functions instead of for loops is called vectorisation

## The apply function

The apply function

- apply(matrix, MARGIN, function)
  - the matrix argument can actually be an array of more than 2 dimenions
  - MARGIN - 1 means vary the rows, 2 means vary the columns
  - function to apply
- e.g. apply(matrix, 1, sum)
  - for each row of the matrix, compute the sum
- e.g. apply(matrix, 2, sum)  
  - for each column of the matrix, compute the mean

## The apply function (code)

```{r}
## Prepare a simple matix
x <- cbind(x1 = 3, x2 = c(3:1, 2:3))
x
# column sum
col.sums1 <- apply(x, 2, sum)
col.sums1
```

## The apply functio(code)

```{r}
# this is equiavlent to
col.sums2 <- NULL
for(i in 1:ncol(x)) {
  col.sums2[colnames(x)[i]] <- sum(x[,i])
}
col.sums2
```

## The other ply functions

In the same family of functions

- lapply(a.list, FUN) 
  - use the FUN on every element in a.list
  - the output is a list 
  
- sapply(a.list, FUN)
  - the s stands for simplify
  - it's the same as lapply, except
  - its output is a vector
  
## The other ply functions (code)

```{r}
x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
# compute the list mean for each list element
lapply(x, mean)
```

## The other ply functions (code)

```{r}
x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
# compute the list mean for each list element
sapply(x, mean)
```

## Take aways

- As much as possible use apply family of functions instead of for/while loops
  - It's faster
  - the more common R paradigm

- On the ply-family of functions
  - http://stackoverflow.com/questions/3505701/r-grouping-functions-sapply-vs-lapply-vs-apply-vs-tapply-vs-by-vs-aggrega
  
## Exercises

- Using the apply function and find out the mean speed and dist in the cars data set
- Do the same with lapply and sapply

## Exercises (Solution)

- Using the apply function and find out the mean speed and dist in the cars data set
- Do the same with lapply and sapply
```{r}
apply(cars, 2, mean)
lapply(cars, mean)
sapply(cars, mean)
```
