---
title: "Lesson 3 - Functions & Debugging"
author: "ZJ"
date: "3 October, 2014"
output: ioslides_presentation
---

## R is a functional language

- A function takes some inputs and produce an output
- Funcitonal languages are programming languages where functions play a key and prominent role
- In R you can pass functions as arguments to functions
  - In a sense functions are treated like any other data
  - this is a hallmark of functional programming languages
  - The most hardcore functional languages don't use loops like **for** and **while**
  - In R you can use for and while loops

## Primitive & user-defined functions

- There are two broad types of functions
  - primitive e.g max, min, sum
  - user defined
```{r}
# primitive function has the
sum
```

## Define your own Function

```{r}
# create a funciton that adds two numbers together
# the output of the function is by default the function's last line
add2 <- function(a,b) {
  a + b
}

# create a function that returns T is a is even
# or one can specify a returned value via the return function
is.even <- function(a) {
  if(a %% 2 ==0) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

## Definte your own function

```{r}
is.even(10)
is.even(99)
```


## User-defined function

You can see the function's source by entering the functions name in the console
```{r}
# see the source code for is.even
is.even
```

## The ... argument {.smaller}

- The ... is only applicable inside functions. It means all other arguments
```{r}
my.func <- function(x, y, z = 2, ...) {
  x2 <- x * x
  y2 <- y * y
  
  other.args <- list(...)
  for(a in other.args) {
    print(paste("printing ", a))
  }
  
  return(x2 + y2 +  z + length(other.args))
}

my.func(1, 2, w = 4, z = 5, a = 8)
```

## The ... argument (cont)

```{r}
my.func(1,2, w = 4, z = 5, a = 8)
```

## Exercise

- Write a function that returns TRUE if passed an odd number
- Write a function that accepts an argument **x** and **...**. It returns the sum of x and ...

## Exercise (Solutions)

- Write a function that returns TRUE if passed an odd number
- Write a function that returns accepts an argument **x** and **...**. It returns the sum of x and ...
```{r}
# Checks if x is odd
is.odd <- function(x) {
  x %% 2 == 1
}

# Sums x with all the elements in ...
addthem <- function(x, ...) {
  args <- list(...)
  for(a in args) {
    x <- x + a
  }
  x
}
```

## Exercise

- Write a piece of code that accepts two vectors, x and y, as input and returns TRUE
  - if all the elements of the first vector is larger than the second vector
  - OR if one and only one element is in the first vector is larger then the second vector

## Exercise (Solution)

```{r}
ans.func <- function(x,y) { # if you already know functions
  if(all(x >y)) {
    return(TRUE)
  } else if (length(which(x > y))) {
    return(TRUE)
  }
  return(FALSE)
}
```

## Infix functions

You can define **infix** functions by surrounding the name with `%%`
```{r}
`%plus%` <- function(a, b) {
  a + b
}

2 %plus% 3
```

## Debugging

RStudio offers a step by step debugging tool

- Insert **browser()** any where you want to stop the code
- or **browser(expr = some.condition )** then the debugging process will only be triggered if some.condition is TRUE
Commands while in debug mode
- n - next line
- c - continue until next break
- Q - quit the debugger

## Exercise

Debug this function
write a a function that check
```{r}
my.func <- function(a,b) {
  a.over.b <- a/b
  mean(a.over.b)
}

my.func(1:3, c(2,0,1))
```

## Exercise (solution)

Debug this function
write a a function that check
```{r}
my.func <- function(a,b) {
  browser()
  a.over.b <- a/b
  mean(a.over.b)
}

my.func(1:3, c(2,0,1))
```

## Exercise 

Debug this function
- iris a famous dataset that comes pre-installed with R
```{r,error=FALSE,eval=FALSE}
# Hint: Type ?subset in the console to look at the definition
# Hint: Use names(iris) 

# compute the average petal width of a particular species
avg_petal_width <- function(Species) {
  iris.subset <- subset(iris$Species == Species) 
  mean(iris.subset$Petal_Width)
}

avg_petal_width("Setosa")
```

## Exercise (Solution)
```{r}
avg_petal_width <- function(Species) {
  iris.subset <- subset(iris, Species == Species)
  mean(iris.subset$Petal.Width)
}

avg_petal_width("setosa")
```

## Function Scoping rules {.small}

- Variables defined outside a function can used inside a function
- Parameters defined as part of the function definition will be used instead if the variable name clashes with one defined outside the function
```{r}
a <- 2 

add2 <- function(a) {
  a + 2
}

add2(1)


add2_no_arg <- function() {
  a +  2
}

add2_no_arg()
```

## Superassignment

- You can assign a new value to a variable from inside a function
- by using the superassignment operator <<- 
```{r}
a <- 2 

change_a <- function() {
  a <<- 3
}

change_a()

a

```

## Functions & Closures (Advanced topic)

- Function is just like any other data
- So a function can also return a function
```{r}
# create a function that starts counting from n
# and each time you run it it increments by one
increment <- function(n) {
  n <- n - 1
  
  increment.fn <- function() {
    n <<- n + 1
    return(n)
  }
  
  return(increment.fn)
}

# create a function that start counting at 8
incr8 <- increment(8)
```

## Function & Closures (cont)

```{r}
incr8
# let's see this a few times
incr8()
incr8()
```


## Recursion - Fibonacci Numbers

```{r}
# returns the nth Fibonacci number
fib <- function(n) {
	if(n %in% c(1,2)) {
		return(1)
	} else {
		return(fib(n - 1) + fib(n - 2))
	}
}
```

## Fibonacci Numbers

```{r}
# returns the nth Fibonacci number
# performance WARNING: this type of recursion is very slow; not a recommended use case
fib2 <- function(n) {
	if(n %in% c(1,2)) {
		return(1)
	} else {
		return(fib(n - 1) + fib(n - 2))
	}
}
```