---
title: "3b - Install packages & package basics"
author: "ZJ"
date: "2/4/2015"
output:
  ioslides_presentation:
    logo: logo.PNG
    transition: faster
---

## Modern data manipulation

- R is continually improving 
- R programmers write packages and publish them for others to use for free
  - Alot of academic papers now publish a R pacakge along with their paper
- number of R packages on CRAN has recently exceeded 6000


## Important Repositories of R packages

* The Comprehensive R Archive Network (CRAN)
  *  The largest repository; easy to install things
  *  cran.rstudio.com
  
- GitHub
  - Easy installation with the devtools package
  - Increasingly popular
  
- Bioconductor (bioconductor.org)
  - Bioscience related package
  - High quality packages; more curated
  
## Installing packages

```{r}
# install a package from CRAN by name
# install.packages("devtools")

# install a package from github
# devtools::install_github("hadley/dplyr")
```

## Package Basics {.smaller}

- Once a package has been installed you can use the **require** or **library** functions to include them
- alternatively you can simply copy built packages into the **library** folder in the path where R is installed.
```{r,eval=FALSE}
# either of the below is sufficient to include the package in your code
library(dplyr) # errors immediately if package is not installed
require(dplyr) # does not raise an error immediately; so you can use it to check if a package exists

# if the package is not installed
if(!require(dplyr)) {
  install.packages("dplyr")
  library(dplyr)
}

```

## Package Basics

- You can access a package's functions via ::
- The form is package::function
- using :: I can run functions and do not need to *library* or *require* the pacakge first

```{r}
# dplyr::filter # this will run the filter function from the dplyr package
```
