---
title: "3b - Install packages & package basics"
author: "ZJ"
date: "2/4/2015"
output:
  ioslides_presentation:
    logo: logo.PNG
    transition: faster
---

## R & packages

- R is continually improving
- The improvements to R often come in the form of **packages**
- A **package** is a set of related R functions and codes
- R programmers write packages and publish them for others to use for free
  - Increasingly, academics now publish R packages along with their papers
- The number of R packages on CRAN has recently exceeded 6000


## Important Repositories of R packages

* The Comprehensive R Archive Network (CRAN)
  *  The largest repository; easy to install things
  *  **install.packages** uses CRAN by default
  *  It does not contain the bleeding edge of the packages
  *  cran.rstudio.com
  
- GitHub
  - Easy installation with the devtools package
  - devtools::install_github
  - Increasingly popular
  - Bleeding edge packages
  
- Bioconductor (bioconductor.org)
  - Bioscience related package
  - High quality packages; more curated
  
## Installing packages

```{r}
# install a package from CRAN by name
# install.packages("devtools")

# install a package from github
# devtools::install_github("user_id/package_repository")
# for example
# devtools::install_github("hadley/dplyr")
```

## Package Basics {.smaller}

- Once a package has been installed you can use the **require** or **library** functions to include them
```{r,eval=FALSE}
# either of the below is sufficient to include the package in your code
# library(dplyr) raises an error immediately if package is not installed
library(dplyr) 
# require(dplyr) returns TRUE if the package is already installed; so you can use it to check so
# Also it does not raise an error immediately if the package is not installed
require(dplyr) 

# if the package is not installed you can use this structure to install it
# and prevent errors
if(!require(dplyr)) {
  install.packages("dplyr")
  library(dplyr)
}

```

## Installing packages from Github

- The devtools package allows you to install packages from Github
- Increasingly the latest bleeding edge versions of packages are hosted on Github
- However the CRAN version is often more stable with less bugs
- The below code will let you install packages from Github 
```{r}
 # install.package(devtools)
 # the below code (if run) will install the dplyr package from hadley which is the user id for Hadley Wickham
 # devtools::install_github("hadley/dplyr")
```

## Brief overview of Github
- Github is a website for hosting source code repositories
- Each repository will have a unique name
- user_id/package_repository
- If another user wants to improve the package_repository he/she can fork the repository
- the forked repository will take from of another_user_id/package_repository
```{r}
 # example usage of install_github
 # devtools::install_github("user_id/package_repository")
```

## Common package issues
You may find it difficult to install packages from CRAN and Github especially in corporate IT environments
- CRAN Issue: If install.packages fails under Windows try running
```{r}
# This will tell R to use the same internet proxy settings as Internet Explorer
# If not R may not be able to access the internet
setInternet2()
```
- Github Issue: If install_github fails the issue could be caused by corporate virus scanners
- You may need to follow the instructions on installing packages manually

## Installing packages manually
- Go to your packages's page
- on CRAN there should be a link to a file called *platform* binaries e.g. Windows binaries
  - if there are multiple binaries you may want to choose the one labelled r-release 
  - it is likely to be the stable version
- In RStudio go to "Tools" -> "Install Packages..."
- In the "Install from:" dropdown choose "Package Archive File (.zip; .tar.gz)" 
- In "Package archive" choose the binaires archive file that you have download 
- Click "Install"
- Alternatively you can simply copy the **binaries** (basically pre-built packages) into the **library** folder in the path where R is installed
- If you are interested in building your own package watch this space

## Package Basics

- You can access a package's functions via ::
- The form is package::function
- using :: I can run functions and do not need to *library* or *require* the package first


```{r}
# dplyr::filter # this will run the filter function from the dplyr package
```
