---
title: "Lesson 3c - data.frame"
author: "ZJ"
date: "3 October, 2014"
output: ioslides_presentation
---

## Reading in stuctured data

- read.csv - allows you to read csv files
- read.xls - from the gdata package allows you to read xls files


```{r, results='hide'}
# getwd() # get the current working directory
# if path is unspecified files input/output goes to the working directory

# write.csv("cars.csv")
# cars1 <- read.csv("cars.csv")


# You can read in Excel spreadsheets too
#require(gdata)
#data.dict <- read.xls("~/Dropbox/AnalytixWare/Data/Kaggle/Data Dictionary.xls")

# view the data
# View1(cars1)

```

## data.farme

- In other languages (e.g. SAS) it call be called dataset, tables
- One of the most important data structures in R
- A data.frame is collection of **equal** length vectors
- **iris** and **cars** are two data.frames that come with R
- **head** function and **tails** function will show you the first & last few rows of the data.frame
```{r}
head(cars)
```

##  View your data.frame
```{r}
View(cars)
```

## attach
You can **attach** a data.frame so that all variable are accessible without $
```{r}
names(cars) # show all the variables in cars
head(cars$speed)

attach(cars)
head(speed)

detach(cars) # removes the variables 
```

## data.farme structrue

- has row.names and by default they are integers
- each column is a vector and can be accessed with $
- e.g. cars$speed
- data.frame is like a list with vector elements of the same length
```{r}
head(cars$speed)
```

## data.farme structrue

- Let data be a data.frame
- Then data[1] would give u the first column of data
- then data[,1] gives the first column as a vector
```{r}
head(cars[1]) # keeps only the first colunm, but is still a data.frame
head(cars[,1]) # keeps the first column as vector
cars[1,] # keeps the first row of data
```

## data.farme structrue

Let rows and columns be vectors of integers then you access the rows columns of a data.frame
- data[rows,]
- data[, columns]
- data[rows, columns]
```{r}
cars[1,] # keeps the first row of data
```

## data.frame subsetting

- Indexing are of the form data.frame[rows, columns]
- data.frame[, columns] means show all rows
- data.frame[rows, ] means show all columns
```{r}
cars[5:10,] # the 5th to 10th row
```

## data.frame index subsetting

```{r}
cars[5:10, 2] # the 5th to 10th row's second vector
cars[5:10, ][2]
```

## Exercise

- What is the average speed?
- Find the row in the cars data.frame with the highest dist

## Exercise (Solution)

```{r}
mean(cars$speed)
cars[cars$dist == max(cars$dist),]
```

## subset function
```{r}
# you can use boolean subsetting
cars[cars$dist < 10,]

# to save some typing you can use the subset function
subset(cars, dist < 10)
```

## data.frame functions

```{r}
nrow(cars) # number of rows
ncol(cars) # number of columns
summary(cars)
```

## data.frame functions

```{r}
names(cars)
head(row.names(cars))
head(rownames(cars))
``` 

## data.frame functions

```{r}
# you can use the columns as vectors 
head(cars$dist / cars$speed)
``` 

## Histogram

```{r}
hist(cars$dist)
``` 

## Density

```{r}
plot(density(cars$dist))
``` 

## Historgram & Density

```{r}
hist(cars$dist,freq=F)
lines(density(cars$dist))
``` 

## Barplot

```{r}
barplot(cars$dist,freq=F)
``` 

## Scatterplot

```{r}
plot(cars)
``` 

## Scatterplot with lines

```{r}
plot(cars, type="b") # b stands for both points and lines
``` 

## Basic Distribution Analysis

```{r}
boxplot(cars)
```

## data.frame - creating new columns

```{r}
cars$dist.over.speed <- cars$dist / cars$speed
head(cars)
```

## data.frame - creating new columns (cont)

```{r}
new.vector <- cars$dist / cars$speed
cars$new.vector <- new.vector
head(cars)
```

## data.frame - removing columns

```{r}
cars$new.vector <- NULL
head(cars)
cars[,3] <- NULL
head(cars)
```

## data.frame - removing columns cont

```{r}
cars$new.vector <- new.vector
head(cars)
cars[,"new.vector"] <- NULL
head(cars)
```

## data.frame - cbind

- cbind - column bind
```{r}
new.cars <- cbind(cars, new.vector)
head(new.cars)
```

## data.frame - rbind

- rbind - row bind
```{r}
cars.head.tails <- rbind(head(cars),tail(cars))
cars.head.tails
```

## Exercises

- Using the cars data.frame
- Create a new column called ratio = dist / speed
- Create a new column ratio2 = ratio^2
- remove the column ratio
- What is the sum of ratio2


## Exercises (Solution)

```{r}
cars$ratio <- cars$speed/cars$dist
cars$ratio2 <- cars$ratio^2
cars$raio <- NULL
sum(cars$ratio)
```

## Renaming

```{r}
cars1 <- cars

names(cars1) # show the variable names

names(cars1)[1] <- "speed1" # change the first variable to speed1


# if you don't know where the variable is you can use which
var.names <- names(cars1)

# use which to locate where the variable is located
index <- which(var.names == "speed1")

names(cars1)[index] <- "speed2"
```