---
title: "Lesson 3b - Practical String Manipulation"
author: "ZJ"
date: "16 October, 2014"
output: ioslides_presentation
---

## Strings

- A string of characters
- Strings can be enclosed in "" or ''
```{r}
a <- "this is a string!"
a
b <- as.character(123)
b
```

## Basic String manipulation - paste cont

- One of the most common string functions is **paste**
- it can concatenate strings together
```{r}
# concanates two strings together with a space in between
paste(c("a","b","c"), c(1,2,3))
```

## Basic String functions

```{r}
# nchar gives the number of characters in a string
nchar("?~!abc")
# substring using substr(string, start, stop) function
substr("abcef", 3, 4)
```

## Letters

```{r}
# a vector of the alphabet
letters

# Capitals
LETTERS
```

## Exercise

- Create this string "a1b2c3...z26"
- Write a function that reverses a string

Exercise (Solution)

- Create this string "a1b2c3...z26"
```{r}
# create the string a1b2c3...z26
paste0(letters,1:26,collapse = "")
```

## Exercise (Solution)

- Write a function that reverses a string (advanced)
```{r}
# a funciton that reverse a string
strReverse <- function(x) {
  x.rev <- ""
  for(i in nchar(x):1) {
    x.rev <- paste0(x.rev, substring(x, i, i))
  }
  x.rev
}

# or more advanced
strReverse <- function(x) {
  sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")
}
```

## String Split

One of the most useuful functions is to be able to break up a string into a vector
```{r}
# split the string into a vector by " "
strsplit("a b c", " ")
# split every character into a vector
strsplit("a b c", NULL)
```

## Exercise

- A palidrome is string that when reverse is the same as the original string. E.g. aabaa is a palindrome while abab is not. Write a function that tests if a string is a palindrome
- **combn(vector,n)** is a function that takes a vector and n and returns all n element combinations of the vector. E.g try combn(c(1:3), 2)
- (Advanced) Find the largest palindrome integer that is the product of two 2-digits integers
  - You may find the **which** function useful

## Exercise (Solution)

```{r}
# an efficient version of string reverse function
strReverse <- function(x) {
  sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")
}

# checks if a string is a palindrome
is.palindrome <- function(s) {
  s.rev <- strReverse(s)
  s == s.rev
}
```

## Exercise (Solution) cont

```{r}
# all possible 2-digit integers
a <- 10:99
# a palinrome cannot end with a 0 so remove all multiples of 10
a <- setdiff(11:99, seq(10,90,by=10))
# create all the combinations of 2-digit numbers
a.pair <- combn(a,2)
# find the product of all possible two digit numbers
a.product <- a.pair[1,] * a.pair[2,]
# convert the product to characters
product.char <- as.character(a.product)
# find all palindrome integers index
w <- which(is.palindrome(product.char))
# return the answer
max(a.product[w])
```