---
title: "6 - XGBoost.rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## XGBoost workshop

### Kaggle competition
Online data science competition

We will work through the GiveMeSomeCredit competition about building credit scoring application
https://www.kaggle.com/c/GiveMeSomeCredit/data

### Read in the data

```{r}
library(data.table)
df = fread("data/cs-training.csv")
```

### Simple Data profiler

```{r data profile, include=FALSE}
library(DataExplorer)
?DataExplorer::create_report(df)
```

## Creating plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
df_column = df[,-"SeriousDlqin2yrs", with = F]
df_label = df[,SeriousDlqin2yrs]

xdm = xgboost::xgb.DMatrix(
  label = df_label, 
  data = as.matrix(df_column))
```

## Subit to kaggle
```{r}
xgm = xgboost::xgboost(
  data=xdm,
  objective = "binary:logitraw",
  nrounds = 2,
  tree_method="exact",)
```

## But how does XGBoost work?

## Why is popular?

## What is Gain?

https://stackoverflow.com/questions/33520460/how-is-xgboost-cover-calculated

## How to learn data science?

For R read r-bloggers.com everyday.

