---
title: "Lesso 2 - Flow Control & Loops"
author: "ZJ"
date: "2 October, 2014"
output: ioslides_presentation
---

## Conditional branching - if

The if statement is perhaps the most ubiquituous programming construct
```{r}
something.is.true <- TRUE
something.else.is.true <- TRUE

if (something.is.true) {
  #do this
} else if(something.else.is.true) {
  #do this other thing
} else {
  #do this is nothing else is true
}
```

## Conditional branching - if example

Compute income tax

```{r}
income <- 50000
if (income <= 20000) {
  tax <- 0
} else if (income <= 30000) {
  tax <- 200 + (income - 20000) * 0.02
} else if (income <= 40000) {
  tax <- 350 + (income - 30000) * 0.035
} else {
  tax <- 42350 + (income - 320000)  * 0.20
}

tax
```

## ifelse

By assigning a value based on the true or false

```{r}
a <- ifelse(TRUE, 1, 2)
a
b <- ifelse(a == 0, 1, 2)
b
```

## ifelse

```{r}
# ifelse works on vectors
v <- 1:5
v1 <- ifelse(v < 3, v, v * 2)
v1
```

## for loop

- If possible in R you should avoid using for loops
- The R programming style prefers vectorization
  - As much as possible AVOID code that deals with each element individual
```{r}
for(i in 1:3) {
  print(i)
}
```

## while loop

```{r}
i <- 0
while(i <= 2) {
  i <- i + 1
  print(i)
}
```

## Exercises

- Print all numbers below 100 skipping any number that is divisible by 2
  - Hint use the %% operation
- Find the prime factors of 24747. E.g. 15 has two factors, 3 and 5. 
  - a prime is a number that is wholly divisible by 1 or itself. The first few primes are 2, 3, 5, 7, 11
- (Advanced) Find the sum of all primes less than 100