---
title: 'Introduction to R and Data Science: Lesson 1'
author: "ZJ"
date: "2 October, 2014"
output: ioslides_presentation
---

## About Me

- Name: Dai Zhuo Jia (ZJ)
- Australian-Chinese
- B.Sc (Pure maths) and Master of Statistics
- Been a useR since 2003
- Author of R package ShinySky (https://github.com/AnalytixWare/ShinySky)
- Worked in Banking (risk analysis) in Australia and Singapore

## What is R?

**R** is 

- a programming *environment*, not just a language
- for doing statistics (2000's - analytics, 2010's data science)
- first developed in New Zealand in 1998 by **R**oss Ihaka and **R**obert Gentleman
- it's actually based on the S language deviced in mid-1970's
- S is for statistics

## Why Learn R?

- Data Science is one of fastest growing fields
- R is one of the premiere languages in this field


## What is R great for?

- Statistical Analysis

```{r}
summary(cars)
```

## What is R great for?
  
- Graphing

```{r, echo=FALSE}
plot(cars)
```

## Starting out
  
- Install R (http://cran.rstudio.com/)
- Install RStudio (http://www.rstudio.com/ide/download/desktop)
- Go to the console
- type 1 + 1, then enter. You get

```{r}
1 + 1
```

## R is interactive
  
- The result of what you type will get reflected right away
- R does not do integer arithmetics, so 3 / 2 does NOT give you 1
```{r}
3 / 2
```


## Arithmetic in R

Operation  | Usage | Example 
------------- | ------------- | -------------
Addition  | a + b | 
Subtraction  | a - b | 
Mulitplication  | a * b | 
Divison  | a / b | 
Exponetiation | a ^ b or a ** b | 
Modulo | a %% b | 5 %% 3 gives 2
Matrix Mult | a %*% b | 



## Variables

You can set values to variables, called **assignment**
```{r}
a = 1
b <- 2
3 -> d
e <- a + b + d
e
```

## Variable Name Conventions

Rule | Usage &  Example
------------- | ------------- 
Can use alphanumeric | a-z A-Z 0-9
. and _ can be used |  var.name & var_name are valid name
Cannot start with numbers or _ | 1name & _name are NOT allowed
Most other symbols (e.g. ,!~`'"-\$ ) can NOT be used | abc! is NOT a valid name 
Case SENSITIVE | abc & ABC are two distinct variable names

                    
## Is this a valid variable name?

- a_
- 1a2
- a1b
- ._
- _abc
- .89

## Disregard all rules if you use backtick (`)

Disregard all rules (except case sensivity) with back-tick \\`
```{r}
`1 whatever name you want !$` <- 100
`1 whatever name you want !$`
```

## Vectors in R

One of most useful and powerful features of R is working with vectors

The main command is **c** whichs stands for **c**ombine

```{r}
primes.1.to.5 <- c(2,3,5,7,11)
sum(primes.1.to.5)
diff(primes.1.to.5)
```

## More Vector Operations

```{r}
primes.1.to.5 <- c(2,3,5,7,11)
max(primes.1.to.5)
min(primes.1.to.5)
```

## Set Operations

```{r}
vector.a <- c(1,2,3)
vector.b <- c(3,4,5)
vector.c <- c(vector.a,vector.b)
vector.c
unique(vector.c)
intersect(vector.a, vector.b)
```

## Sorting

```{r}
vector.c.desc <- sort(vector.c, decreasing = TRUE)
vector.c.desc
sort(vector.c.desc)
# or simply return the order of the values
order(vector.c.desc)
```

## Exercise



## Arithmetic Operations on vectors

Every element in the vector happens
```{r}
numbers3 <- 1:5
numbers3 * 2
```

## What happens internally
- R needs two vectors of the SAME length
- so 2 gets converted to c(2,2,2,2,2)
- then c(1,2,3,4,5) * c(2,2,2) = c(1 * 2, 2 * 2, 3 * 2,4 *2, 5 *2)
- gives c(2,4,6,8,10)

## Vector Recycling

```{r}
numbers3 <- 1:5
numbers3 * c(1,2)
```

## Can you understand why?
- R needs two vectors of the SAME length
- so it **recycles** the shorter vector
- so c(1, 2) gets repeated up to c(1, 2, 1, 2, 1)
- c(1,2,3,4,5) * c(1,2,1,2,1) = c(1 * 2, 2 * 2, 3 * 1, 4 * 2, 5 * 1)
- gives c(1, 4,3 ,8 ,5)

## Exercise

What is the output of 
```{r, eval=FALSE}
sum(2:5)
diff(10:15)
2:5 - 1
1:5 * 1:3
```

## Exercise (Solution) {.smaller}

```{r}
sum(2:5)
diff(10:15)
2:5 - 1
1:5 * 1:3
```

## Number Indexing {.smaller}

```{r}
numbers10 <- 1:5
numbers10[2]
numbers10[2:4]
numbers10[4:7]
# you can also use indexing to change the data to this
numbers10[2] <- 100
numbers10
numbers10[c(1,3:4)] <- c(999, 999, 999)
numbers10
```

## Number Indexing - removal

```{r}
# you can remove the i-th element by doing this
i = 4
numbers10[-i] # this will remove the 4th element
# however numbers10 is actually not changed 
numbers10
# to change the content of numbers10
numbers10 <- numbers10[-i]
```

## Exercises

- Create a vector (5, ...., 2, 1, 0, 1, 2,...,5)
- Remove the 0 element from the vector

## Exercises (Solution)

```{r}
x <- c(5:0,1:5)
x
x1 <- x[-6]
x1
```

## Boolean

Boolean is the TRUE/FALSE data type

- TRUE or T
- FALSE or T
- == tests for equality e.g. 1 == 1 is TRUE 0 == 2 is FALSE
- 2 > 1 evaluates to TRUE
- 2 < 3 evaluates to FALSE

## Boolean (TRUE/FALSE) Indexing

```{r}
numbers10 <- 1:5
numbers10[c(T, F, T, F, T)]
```

## Filtering

Find all multiples of 2 less than or equal to 10
```{r}
numbers10 <- 1:10
numbers10[c(FALSE,TRUE)] # recycling
numbers10 %% 2 == 0
numbers10[numbers10 %% 2 == 0]
```

## Exercise

- Store all the multiple of 3s under 100 in a vector
- Store all the multiple of 5s less than or equal to 100 in a vector
- Find those numbers that are either multiples of 3 and/or multiples of 5, no repeats allowed
- Find their sum

## Exercise (Solution)

```{r}
# solution 1
multiple.3 <- (1:100)[1:100 %% 3 == 0]
multiple.5 <- (1:100)[1:100 %% 5 == 0]
final.ans <- unique(c(multiple.3, multiple.5))
sum(final.ans)
# solution 2
final.ans <- (1:100)[1:100 %% 3 == 0 | 1:100 %% 5 == 0]
sum(final.ans)
```

## Boolean Operations

Name  | R Symbol | Example
------------- | ------------- | -------------
And  | & | 1 == 1 & 0 == 0
Or  | &#124;  | T &#124; F
Not | ! | 0 != 0, !1 == 1

## Truth Table | and (&)

a  | b | a & b
------------- | ------------- | -------------
FALSE | FALSE | FALSE
FALSE | TRUE | FALSE
TRUE | FALSE | FALSE
TRUE | TRUE | TRUE

## Truth Table | or (|)

a  | b | a &#124; b
------------- | ------------- | -------------
FALSE | FALSE | FALSE
FALSE | TRUE | TRUE
TRUE | FALSE | TRUE
TRUE | TRUE | TRUE

## Exercise

Find the sum of all numbers below 100 that are a multiple of 3 but NOT a multiple of 5

## Exercise (Solution)

Find the sum of all numbers below 100 that are a multiple of 3 but NOT a multiple of 5
```{r}
numbers100 <- 1:100
final <- numbers100[numbers100 %% 3 == 0 & !numbers100 %% 5 ==0]
sum(final)
```

## More useful Boolean functions

```{r}
# returns true if ANY element of the argument vector is TRUE
any(c(F,T,F))
# returns true if ALL element of the argument vector is TRUE
all(c(F,T,F))
all(c(T,T,T))
```

## More useful vector functions - which family

The **which** family of functions is extremely useful
- **which** accepts a vector of boolean and returns the index of the true position
- **which.min** & **which.max** returns the index of the FIRST minimum/maximum value
```{r}
which(1:10 > 5)

which.min(c(1:3,0,0,4:6))
```

## Exercise

- Write a function that accepts two vectors as input and returns TRUE
  - if all the elements of the first vector is larger than the second vector
  - OR if one and only one element is in the first vector is larger then the second vector

## Exercise (Solution)

```{r}
ans.func <- function(x,y) {
  if(all(x >y)) {
    return(TRUE)
  } else if (length(which(x > y))) {
    return(TRUE)
  }
  return(FALSE)
}
```

## Vector's elements must be of the same type

You can have vectors of numbers, boolean, and strings (characters)
```{r}
c(1,2,3)
c(T,F,T)
c(10,T,F,T,100)
```

## Vector's elements must be of the same type

If you mix the types it will converted to first, strings, and second, numerics.
```{r}
c("a","b","c")
c(1,"a",T)
```

## Seeking Help

The documentation will show in the **Help** tab
```{r ,eval=FALSE}
?sum
??deeper_search

# getting examples
example(min)
```

## Tips

- The power of TAB: type the first few characters of a function then press tab
- stackoverflow.com General programmings questions